---
title: "The Ames Housing Data"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
data(ames, package = "modeldata")
dim(ames)
```

# 1 Exploring features of homes in Ames

```{r}
library(tidymodels)
tidymodels_prefer()
```

```{r}
ggplot(ames, aes(x = Sale_Price)) +
  geom_histogram(bins = 50, col = "white")
```

Log-transformation:

-   no houses will be predicted with negative ale prices

-   errors in predicting expensive houses will not have an undue influence on the model

-   from a statistical point of view, a logarithmic transform may also stabilize the variance in a way that makes inference more legitimate

```{r}
ggplot(ames, aes(x = Sale_Price)) +
  geom_histogram(bins = 50, col = "white") +
  scale_x_log10()
```

The units of the model coefficients might be more difficult to interpret, as will measures of performance (e.g. the residuals will also be on the log scale)

```{r}
ames <- ames |> 
  mutate(Sale_Price = log10(Sale_Price))
```

Spatial inormation:

```{r}
ggplot(ames, aes(x = Longitude, y = Latitude,
                 col = Neighborhood)) +
  geom_point() +
  theme(legend.position = "bottom")
```

Some basic questions that could be examined during this exploratory stage include:

-   is there anything odd or noticeable about the distributions of the individual predictors? Is there much skewness or any pathological distributions?

-   are there high correlations between predictors? For example, there are multiple predictors related to house size. Are some redundant?

-   are there associations between predictors and the outcomes?

# 2 Chapter summary

```{r}
library(tidymodels)
data(ames)
ames <- ames |> mutate(Sale_Price = log10(Sale_Price))
```
