---
title: "A model workflow"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

-   Using a workflow concept encourages good methodology since it is a single point of entry to the estimation components of a data analysis

-   It enables the user to better organize projects

# 1 Where does the model begin and end?

Linear regression produces a model equation of

$$
\hat{y_i}=\hat{\beta_0}+\hat{\beta_1}x_{i1}+...+\hat{\beta_p}x_{ip}
$$

While this is a linear model, it is linear only in the parameters. The predictors could be nonlinear terms (e.g. $log(x_i)$).

A variety of choices and additional steps often occur before the model is fit:

-   exclusion of some of the predictors from the analysis

-   imputation of an important missing predictor

-   transformation of the scale of a predictor

There may also be operations that occur after the model is created (e.g. cutoff threshold for a classification model).

# 2 Workflow basics

```{r}
library(tidymodels)
tidymodels_prefer()
```

```{r}
lm_model <- 
  linear_reg() |> 
  set_engine("lm")
```

A workflow always requires a `parsnip` model object:

```{r}
lm_wflow <- 
  workflow() |> 
  add_model(lm_model)

lm_wflow
```

A standard R formula can be used as a preprocessor:

```{r}
lm_wflow <- 
  lm_wflow |> 
  add_formula(Sale_Price ~ Longitude + Latitude)

lm_wflow
```

Workflows have a `fit()` method that can be used to create the model.

```{r}
lm_fit <- fit(lm_wflow, ames_train)
lm_fit
```

We can also `predict()` on the fitted workflow:

```{r}
predict(lm_fit, ames_test |> slice(1:3))
```

Both the model and preprocessor can be removed or updated:

```{r}
lm_fit |> update_formula(Sale_Price ~ Longitude)
```

# 3 Adding raw variables to the `workflow()`

```{r}
lm_wflow <- 
  lm_wflow |> 
  remove_formula() |> 
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))
lm_wflow
```

```{r}
fit(lm_wflow, ames_train)
```

# 4 How does a `workflow()` use the formula?

## 4.1 Special formulas and inline functions

```{r}
library(lme4)
lmer(distance ~ Sex + (age | Subject), data = nlme::Orthodont)
```

Standard R methods can't properly process this formula:

```{r}
model.matrix(distance ~ Sex + (age | Subject), data = nlme::Orthodont)
```

The special formula has to be processed by the underlying package code, not the standard `model.matrix()` approach.

```{r}
library(multilevelmod)
```

```{r}
multilevel_spec <- linear_reg() |> set_engine("lmer")

multilevel_workflow <- 
  workflow() |> 
  # pass the data along as-is:
  add_variables(outcome = distance, predictors = c(Sex, age, Subject)) |> 
  add_model(multilevel_spec,
            # This formula is given to the model
            formula = distance ~ Sex + (age | Subject))

multilevel_fit <- fit(multilevel_workflow, data = nlme::Orthodont)

multilevel_fit
```

We can use the previously `strata()` function from the `survival` package for survival analysis:

```{r}
library(censored)
```

```{r}
parametric_spec <- survival_reg()

parametric_workflow <- 
  workflow() |> 
  add_variables(outcome = c(fustat, futime), predictors = c(age, rx)) |> 
  add_model(parametric_spec,
            formula = Surv(futime, fustat) ~ age + strata(rx))

parametric_fit <- fit(parametric_workflow, data = ovarian)
parametric_fit
```

# 5 Creating multiple workflows at once
