---
title: "Feature Engineering with recipes"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# 8.1 A simple `recipe()` for the Ames housing data

```{r}
lm(Sale_Price ~ Neighborhood + log10(Gr_Liv_Area) + Year_Built + Bldg_Type, data = ames)
```

```{r}
library(tidymodels)
tidymodels_prefer()
```

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_dummy(all_nominal_predictors())

simple_ames
```

# 8.2 Using recipes

```{r}
#| error: true
lm_wflow |> 
  add_recipe(simple_ames)
```

```{r}
lm_wflow <- 
  lm_wflow |> 
  remove_variables() |> 
  add_recipe(simple_ames)

lm_wflow
```

```{r}
lm_fit <- fit(lm_wflow, ames_train)
```

```{r}
predict(lm_fit, ames_test |> slice(1:3))
```

```{r}
# Get the recipe after it has been estimated:
lm_fit |> 
  extract_recipe(estimated = TRUE)
```

```{r}
# To tidy the model fit:
lm_fit |> 
  # This returns the parsnip object:
  extract_fit_parsnip() |> 
  # Now tidy the linear model object:
  tidy() |> 
  slice(1:5)
```

# 8.3 How data are used by the `recipe()`

# 8.4 Examples of recipe steps

## 8.4.1 Encoding qualitative data in a numeric format

```{r}
ggplot(ames_train) +
  geom_bar(aes(y = Neighborhood))
```

To lump the bottom 1% of the neighborhoods into a new level called "other":

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.01) |> 
  step_dummy(all_nominal_predictors())
```

```{r}
simple_ames
```

## 8.4.2 Interaction terms

```{r}
ggplot(ames_train, aes(x = Gr_Liv_Area, y = 10^Sale_Price)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~ Bldg_Type) + 
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, 
              color = "lightblue") + 
  scale_x_log10() + 
  scale_y_log10() + 
  labs(x = "Gross Living Area", y = "Sale Price (USD)")
```

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.01) |> 
  step_dummy(all_nominal_predictors()) |> 
  # Gr_Liv_Area is on the log scale from a previous step
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_"))
```

## 8.4.3 Spline functions

Splines replace the existing numeric predictor with a set of columns that allow a model to emulate a flexible, nonlinear relationship. As more spline terms are added to the data, the capacity to nonlinearly represent the relationship increases.

```{r}
library(patchwork)
library(splines)
```

```{r}
plot_smoother <- function(deg_free) {
  ggplot(ames_train, aes(x = Latitude, y = 10^Sale_Price)) +
    geom_point(alpha = 0.2) +
    scale_y_log10() +
    geom_smooth(
      method = lm,
      formula = y ~ ns(x, df = deg_free),
      color = "lightblue",
      se = FALSE
    ) +
    labs(title = paste(deg_free, "Spline Terms"),
         y = "Sale Price (USD")
}
```

```{r}
( plot_smoother(2) + plot_smoother(5) ) / ( plot_smoother(20) + plot_smoother(100))
```

```{r}
recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude, data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.01) |> 
  step_dummy(all_nominal_predictors()) |> 
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) |> 
  step_ns(Latitude, deg_free = 20)
```

## 8.4.4 Feature extraction

PCA is a linear extraction method, meaning that each new feature is a linear combination of the original predictors.

## 8.4.5 Row sampling method

-   *Downsampling*: the data keeps the minority class and takes a random sample of the majority class so that class frequencies are balanced

-   *Upsampling* replicated samples from the minority class to balance the classes.

## 8.4.6 General transformations

## 8.4.7 Natural language processing

# 8.5 Skipping steps for new data

# 8.6 Tidy a `recipe()`

```{r}
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude + Longitude, data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.01) |> 
  step_dummy(all_nominal_predictors()) |> 
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) |> 
  step_ns(Latitude, Longitude, deg_free = 20)
```

```{r}
tidy(ames_rec)
```

```{r}
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude + Longitude, data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.01, id = "my_id") |> 
  step_dummy(all_nominal_predictors()) |> 
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) |> 
  step_ns(Latitude, Longitude, deg_free = 20)
```

```{r}
lm_wflow <- 
  workflow() |> 
  add_model(lm_model) |> 
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)
```

```{r}
estimated_recipe <- 
  lm_fit |> 
  extract_recipe(estimated = TRUE)
```

```{r}
tidy(estimated_recipe, id = "my_id")
```

```{r}
tidy(estimated_recipe, number = 2)
```

# 8.7 Column roles

```{r}
#| error: true
ames_rec |> 
  update_role(address, new_role = "street address")
```

# 8.8 Chapter summary

```{r}
library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.8, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)

ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude + Longitude, data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.01) |> 
  step_dummy(all_nominal_predictors()) |> 
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) |> 
  step_ns(Latitude, Longitude, deg_free = 20)

lm_model <- linear_reg() |> set_engine("lm")

lm_wflow <- 
  workflow() |> 
  add_model(lm_model) |> 
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)
```
