---
title: "Judging Model Effectiveness"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# 9.1 Performance metrics and inference

# 9.2 Regression metrics

```{r}
ames_test_res <- predict(lm_fit, 
                         new_data = ames_test |> select(-Sale_Price))
```

```{r}
ames_test_res
```

```{r}
ames_test_res <- bind_cols(ames_test_res, 
                           ames_test |> select(Sale_Price))
```

```{r}
ames_test_res
```

```{r}
ggplot(ames_test_res, aes(x = Sale_Price, y = .pred)) +
  # Create a diagonal line:
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(y = "Predicted Sale Price (log10)",
       x = "Sale Price (log10)") +
  # Scale and size the x- and y-axis uniformly:
  coord_obs_pred()
```

Compute the root mean squared error for this model:

```{r}
rmse(ames_test_res, truth = Sale_Price, estimate = .pred)
```

Add r squared and the mean absolute error:

```{r}
ames_metrics <- metric_set(rmse, rsq, mae)
ames_metrics(ames_test_res, truth = Sale_Price, estimate = .pred)
```

# 9.3 Binary classification metrics

```{r}
data("two_class_example")
tibble(two_class_example)
```

```{r}
# A confusion matrix:
conf_mat(two_class_example, truth = truth, estimate = predicted)
```

Accuracy: (TP + TN) / (TP + TN + FP + FN)

```{r}
(227 + 192) / (227 + 192 + 50 + 31)
```

```{r}
# Accuracy:
accuracy(two_class_example, truth, predicted)
```

Matthews correlation coefficient: \[(TP \* TN) - (FP \* FN)\] / sqrt\[(TP+FP)(TP+FN)(TN+FN)\]

```{r}
((227 * 192) - (31 * 50)) / sqrt((227 + 31) * (227 + 50) * (192 + 50))
```

```{r}
# Matthews correlation coefficient:
mcc(two_class_example, truth, predicted)
```

```{r}
# F1 metric:
f_meas(two_class_example, truth, predicted)
```

Combining these three classification metrics together:

```{r}
classification_metrics <- metric_set(accuracy, mcc, f_meas)
classification_metrics(two_class_example, 
                       truth = truth, estimate = predicted)
```

Choose the second level as the event:

```{r}
f_meas(two_class_example, truth, predicted, event_level = "second")
```

```{r}
two_class_curve <- roc_curve(two_class_example, truth, Class1)
two_class_curve
```

```{r}
roc_auc(two_class_example, truth, Class1)
```

```{r}
autoplot(two_class_curve)
```

# 9.4 Multiclass classification metrics

```{r}
data("hpc_cv")
tibble(hpc_cv)
```

```{r}
accuracy(hpc_cv, obs, pred)
mcc(hpc_cv, obs, pred)
```

```{r}
class_totals <- 
  count(hpc_cv, obs, name = "totals") |> 
  mutate(class_wts = totals / sum(totals))
class_totals
```

```{r}
cell_counts <- 
  hpc_cv |> 
  group_by(obs, pred) |> 
  count() |> 
  ungroup()
```

```{r}
# Compute the four sensitivities using 1-vs-all
one_versus_all <- 
  cell_counts |> 
  filter(obs == pred) |> 
  full_join(class_totals, by = "obs") |> 
  mutate(sens = n / totals)
one_versus_all
```

```{r}
# Three different estimates:
one_versus_all |> 
  summarise(
    macro = mean(sens),
    macro_wts = weighted.mean(sens, class_wts),
    micro = sum(n) / sum(totals)
  )
```

```{r}
sensitivity(hpc_cv, obs, pred, estimator = "macro")
sensitivity(hpc_cv, obs, pred, estimator = "macro_weighted")
sensitivity(hpc_cv, obs, pred, estimator = "micro")
```

```{r}
roc_auc(hpc_cv, obs, VF, F, M, L)
```

```{r}
roc_auc(hpc_cv, obs, VF, F, M, L, estimator = "macro_weighted")
```

```{r}
hpc_cv |> 
  group_by(Resample) |> 
  accuracy(obs, pred)
```

```{r}
# Four 1-vs-all ROC curves for each fold
hpc_cv |> 
  group_by(Resample) |> 
  roc_curve(obs, VF, F, M, L) |> 
  autoplot()
```

# 9.5 Chapter summary
