---
title: "Resampling for evaluating performance"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# 10.1 The resubstitution approach

We can fit a random forest model to the training set:

```{r}
rf_model <- 
  rand_forest(trees = 1000) |> 
  set_engine("ranger") |> 
  set_mode("regression")

rf_wflow <- 
  workflow() |> 
  add_formula(
    Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type +
      Latitude + Longitude) |> 
  add_model(rf_model)

rf_fit <- rf_wflow |> 
  fit(data = ames_train)
```

How should we compare the linear and random forest models?

We predict the training set to produce what is known as an *apparent metric* or *resubstitution metric*. This function creates predictions and formats the results:

```{r}
estimate_perf <- function(model, dat) {
  # Capture the names of the `model` and `dat` objects
  cl <- match.call()
  obj_name <- as.character(cl$model)
  data_name <- as.character(cl$dat)
  data_name <- gsub("ames_", "", data_name)
  
  # Estimate these metrics:
  reg_metrics <- metric_set(rmse, rsq)
  
  model |> 
    predict(dat) |> 
    bind_cols(dat |> select(Sale_Price)) |> 
    reg_metrics(Sale_Price, .pred) |> 
    select(-.estimator) |> 
    mutate(object = obj_name, data = data_name)
}
```

The resubstitution statistics are:

```{r}
estimate_perf(rf_fit, ames_train)
```

```{r}
estimate_perf(lm_fit, ames_train)
```

Apply the random forest model to the test set for final verification:

```{r}
estimate_perf(rf_fit, ames_test)
```

The test set RMSE estimate, 0.0701, is much worse than the training set value of 0.0364!

Many predictive models are capable of learning complex trends from the data. In statistics, these are commonly referred to as *low bias models*.

# 10.2 Resampling methods
