---
title: "Comparing models with resampling"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# 11.1 Creating multiple models with workflow sets

```{r}
basic_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude + Longitude, data = ames_train) |> 
  step_log(Gr_Liv_Area, base = 10) |> 
  step_other(Neighborhood, threshold = 0.1) |> 
  step_dummy(all_nominal_predictors())

interaction_rec <- 
  basic_rec |> 
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_"))

spline_rec <- 
  interaction_rec |> 
  step_ns(Latitude, Longitude, deg_free = 50)

preproc <- 
  list(basic = basic_rec,
       interact = interaction_rec,
       splines = spline_rec)

lm_models <- workflow_set(preproc, list(lm = linear_reg()), cross = FALSE)

lm_models
```

```{r}
lm_models <- 
  lm_models |> 
  workflow_map("fit_resamples",
               # Options to `workflow_map()`:
               seed = 1001, verbose = TRUE,
               # Options to`fit_resamples()`:
               resamples = ames_folds, control = keep_pred)
lm_models
```

```{r}
collect_metrics(lm_models) |> 
  filter(.metric == "rmse")
```

Add the random forest model:

```{r}
four_models <- 
  as_workflow_set(random_forest = rf_res) |> 
  bind_rows(lm_models)
four_models
```

```{r}
library(ggrepel)
```

```{r}
autoplot(four_models, metric = "rsq") +
  geom_text_repel(aes(label = wflow_id), nudge_x = 1/8, nudge_y = 1/100) +
  theme(legend.position = "none")
```

# 11.2 Comparing resampled performance statistics

We gather the individual resampling statistics for the linear models and the random forest.

```{r}
rsq_indiv_estimates <- 
  collect_metrics(four_models, summarize = FALSE) |> 
  filter(.metric == "rsq")
```

```{r}
rsq_wider <- 
  rsq_indiv_estimates |> 
  select(wflow_id, .estimate, id) |> 
  pivot_wider(id_cols = "id", names_from = "wflow_id", 
              values_from = ".estimate")
```

```{r}
corrr::correlate(rsq_wider |> select(-id), quiet = TRUE)
```

Across models, there are large within-resample correlations.

```{r}
rsq_indiv_estimates |> 
  mutate(wflow_id = reorder(wflow_id, .estimate)) |> 
  ggplot(aes(x = wflow_id, y = .estimate, group = id, color = id)) +
  geom_line(alpha = 0.5, linewidth = 1.25) +
  theme(legend.position = "none")
```

A statistical test for the correlations evaluates whether the magnitudes of these correlations are not simply noise:

```{r}
rsq_wider |> 
  with(cor.test(basic_lm, splines_lm)) |> 
  tidy() |> 
  select(estimate, starts_with("conf"))
```

# 11.3 Simple hypothesis testing methods
